Haskell in Computational Biology

### 2013-05-06 jfp.tex
added a note about the first delta in the software.  it sucks.

>>>>

\subsection{Software deltas}

\subsubsection{Reducing cases in Viterbi}

Original Viterbi took a state and two indices as argument, and it
dealt with impermissible transitions by assigning a very large score
(which translates to a very small probability).
Indices 1, 0, and $-1$ were all used to indicate special conditions,
and 
the function had 25~cases overall.

\newcommand\pv{\textit{Predecessor Viterbi}}
\newcommand\ov{{Original Viterbi}}

Our first step in adapting the code was to create \pv.
Like \ov, \pv\ takes a state and two indices as an argument.
But at top level, there is no case analysis on the state, and the only
indices considered are zero and nonzero, so there are only four cases.
(The state is used in two ways: to compute transition probabilities,
and to decide which states may permissibly precede the current state.
The set of permissible preceding states is looked up in a data
structure, not computed by case analysis.)

Here is the general case in the new code:
\smallverbatiminput{hoviterbi}
THERE IS A GRATUITOUS CHANGE: THE TYPE OF ARRAY USED IS DIFFERENT.
WE SHOULD PROBABLY FIX THAT.
In the old code, there are three corresponding cases: one each for
match, insert, and delete states.  
Here is the code for the match case:
\smallverbatiminput{viterbi}
The other case are similar; the second argument to \texttt{aScore} is
always equal to the first (state) argument to \texttt{vee'},
and the argument to \texttt{map from} is different in each case.
(In the new code, this argument is identical to \texttt{preceders !
stateRight}.)
NORMAN THINKS THIS COMPARISON DOESN'T WORK AT ALL :-(

The main benefit is that the Viterbi function is reduced from 90~lines
of code to just 40~lines.
This change laid the groundwork for more experiments.

BENCHMARK RESULTS WERE WILDLY INCONSISTENT.
